### 2011-06-17 thesis/novelty-intro.tex
Changes on related work of novelty detection chapter

\subsection{Review}
Many approaches have been made based on statistical methods. Those often model
the training data with statistic distributions and then perform hypothesis tests.
For example: testing whether two distributions are the same or using distance to
mean value or quartiles.
In some sense all they define a distance measure to normality and employ a
threshold on that.

A common approach is to use density estimation and use the expected probability
of a sample to trigger the sample as novel. Examples are usage of
Gaussian Mixture Models and Parzen-window estimators. In order to be effective
those use data as close as possible to the input features and use
dimension-reduction techniques such as \gls{PCA} to make density
estimation feasible. Note that as dimension increases an exponential number
of data samples would be required to approach the density with the same quality.

\cite{bishop1994novelty} uses that approach by employing a Parzen-window to
estimate the density of the training data on a given input feed into a
neural-network. By triggering on the calculated density they detect samples
that differ from the training data and consider their neural-network output
to be unreliable as the samples are distinct from what the network was trained
with.

A slightly different approach is done by one-class \gls{SVM} approaches that
try to distinguish novelty by separating the training set from all the other
points in the input space. They try to achieve that by enclosing the training
set by some structure (e.g.\ an hyper-sphere)~\cite{bennett2000support}.
Those approaches have been made in line with Vapnik principle of not solving
something hard. As although having access to a perfect probability distribution
of the input would solve the problem, creating such a function is harder than
simply creating a boundary between known data and novel data
\cite{scholkopf2000support}.

Error reconstruction methods have also been used for novelty detection.
They use the assumption that the class to be defined lies on a manifold embedded
on an input space of higher dimensions. By using dimensional reduction
techniques, they try to defined that manifold and calculate how far a new sample
is from it.
One of the most common methods used for that is
\gls{K-PCA}~\cite{scholkopf1997kernel}, which uses the kernel-trick to extend
\gls{PCA} and perform a nonlinear dimensionality reduction on the input.
This technique has been successfully used in novelty detection by
\cite{Hoffmann2007863}.

\subsection{Related Work}
\cite{zhang2005probabilistic} present a probabilistic semantic approach to image
annotation and retrieval with a multi-modal information. They introduce an
hidden variable modelling a concept present on a scene that connects together
information coming from both a bag of words and visual features extracted. By
training that hidden-variable their system becomes aware of possible existent
semantic categories and their relation between both textual and visual features.
Although not able to perform novelty detection, their approach is interesting by
the use of an hidden variable to probabilistic model the multi-modal sensors.

\cite{boutell2006factor} presents a method to perform scene-classification
from low-level region detectors using factor graphs. Those low-level detectors
provide semantic (high-level) cues on the higher level scene-classification
task. Due to the nature of the low-level detectors, they ought to be treated
in a probabilistic way. For that their method uses factor graphs. The method
also exploits spatial relation information between regions to improve the
classification performance. A presented scheme using only pairwise relations
between regions is show to have better performance than try to approach the
correct scheme with a big single-factor. I.e.\ the pairwise relations can be
approximated with the modest amounts of training data, where the single-factor
demands an impractical amount of training data.
No approach is made on it at novelty detection either of regions or scene
categories.

\cite{bishop1994novelty} works shows that assuming a uniform unconditional
distribution novelty detection can be performed by triggering on the
density probability of the training data. The system is employed to detect novel
inputs for a neural-network.

@@

>>>>

\subsection{Review of Novelty Detection Methods}
Many approaches have been made based on statistical methods. Those often model
the training data with statistic distributions and then perform hypothesis tests.
For example: testing whether two distributions are the same or using distance to
mean value or quartiles.
In some sense all they define a distance measure to normality and employ a
threshold on that.

A common approach is to use density estimation and use the expected probability
of a sample to classify the sample as novel. Examples of such techniques are
Gaussian Mixture Models and Parzen-window estimators. In order to be effective,
those rely on data as close as possible to the input features and use
dimensionality\hyp{}reduction techniques such as \gls{PCA} to make density
estimation feasible. Note that as dimension increases an exponential number
of data samples would be required to approach the density with the same quality.

\cite{bishop1994novelty} uses that approach by employing a Parzen-window to
estimate the density of the training data on a given input fed into a
neural-network. Using the calculated density, they detect samples
that differ from the training data and consider their neural-network output
to be unreliable as the samples are distinct from what the network was trained
with.

A slightly different approach is applied in case of by one-class \gls{SVM} approaches that
try to distinguish novelty by separating the training set from all the other
points in the input space. They try to achieve that by enclosing the training
set by some structure (e.g.\ an hyper-sphere)~\cite{bennett2000support}.
Those approaches have been made in line with Vapnik principle of not solving
something hard. As although having access to a perfect probability distribution
of the input would solve the problem, creating such a function is harder than
simply creating a boundary between known data and novel data
\cite{scholkopf2000support}.

Error reconstruction methods have also been used for novelty detection.
They use the assumption that the class to be defined lies on a manifold embedded
into a sample space of higher dimensionality. By using dimensionality\hyp{}reduction
techniques, a manifold is defined and the distance between the manifold and the new sample
is calculated. One of the most common methods used for that purpose is
\gls{K-PCA}~\cite{scholkopf1997kernel}, which uses the kernel-trick to extend
\gls{PCA} and perform a nonlinear dimensionality reduction of the input.
This technique has been successfully used for novelty detection in \cite{Hoffmann2007863}.

\cite{japkowicz1995novelty} also follows a similar approach: \emph{Redundancy
Compression and Non-Redundancy Differentiation} which is a process believed
to happen in the hippocampus. They introduce an auto-encoder
that learns to encode a sample on a considerable smaller description and later reconstructs
the original sample from this smaller description. Their suggested system
learns how to discard and compress redundant information while still be able to recover
the original sample. By training it with samples from a given class, it
is expected that it will badly reconstruct a sample from a different one.

@@

\cite{boutell2006factor} presents a method to perform scene-classification
from low-level region detectors using probabilistic graphical models.
The information obtained by the region detectors is exploited together with the
spatial relations between on the higher level scene-classification task.
A presented scheme using only pairwise relations between regions is show to
have better performance than the correct approach of modelling all regions connectivity
with a large single-factor as the pairwise relations can be
approximated with modest amounts of training data, where the single-factor
demands an impractical amount of training data.
No approach is made on it at novelty detection either of regions or scene
categories.
